---
title: "Win your office pool with mRchmadness"
author: "Eli Shayer and Scott Powers"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Win your office pool with mRchmadness}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(mRchmadness)
```

It's Selection Sunday! The full field of 68 teams for the NCAA men's basketball
tournament has been announced, and now that colleague of yours who cares way
too much about sports is inviting you to submit a bracket to the office pool.
You have a few days to choose one of the 9,233,372,036,854,775,808 possible
brackets to call your own. But which one?

### Scraping data

Is this the first day you've paid any attention to college basketball all year?
We've got you covered, with scraping functions that will tell you everything
that went down this season. You can scrape all of the game scores from the
2016-2017 season using the **scrape.game.results** function. In fact, you can
use this same function to scrape game results for any other season, past or
future, as long as the data are available on ESPN.com. Later, we'll use the
data from the current year to predict matchup results.

```{r, eval = FALSE}
games.2017 = scrape.game.results(2017)
```

It also helps to know who everybody else is picking to advance in each round of
the tournament, so that you know what you need to beat. Another function,
**scrape.population.distribution**, grabs these data for you, based on the
population distribution of picks according to ESPN. Note that these data are
currently available online only for the 2016 and 2017 tournaments.

```{r, eval = FALSE}
pred.pop.2017 = scrape.population.distribution(2017)
```

We've made life even easier for you by pre-scraping **games.men.2017**,
**pred.pop.men.2017**, **games.men.2016** and **pred.pop.men.2017**. These are
all available as datasets in the package. You're welcome.

```{r}
head(games.men.2017)
head(pred.pop.men.2017)
```

The game results are stored using team IDs, and we will see that so is the
tournament bracket. But you can use **teams.men** to see who these teams are.

```{r}
head(teams.men)
```

### Predicting matchup results

Now that you know everything that happened this year, you know how good all of
the teams are, right? The Bradley-Terry model can help you make sense of these
scores. According to this model, in game $i$ between home team $H_i$ and away
team $A_i$, the number of points $y_i$ (which may be negative) by which the
home team wins has the following distribution:
$$y_i\sim\mathcal N(\beta_{H_i} - \beta_{A_i}, \sigma^2)$$
where the $\beta$'s represent the unknown quality of the teams. We provide the
**bradley.terry** function to estimate the $\beta$'s, $\sigma$ and the
corresponding probabilities for each team beating each other team.  This
function returns a matrix of probabilities, with one row for each team and one
column for each team. Each entry of the matrix gives the estimated probability
of the team in that row beating the team in that column. Using this, we can
estimate for the probability that North Carolina beats Duke, for example.

```{r}
set.seed(1)
prob.matrix = bradley.terry(games = games.men.2017)
prob.matrix["153", "150"]
```

Not satisfied with this model? Perhaps you'd trust the predictions of the
pundits at FiveThirtyEight.com more. We've provided those in a data.frame as
well.

```{r}
head(pred.538.men.2017)
```

### Simulating the tournament

We've already taken care of those pesky first four games for you and stored the
bracket in a vector called **bracket.men.2017**. This lists the tournament
teams in order of overall seed.

```{r}
head(bracket.men.2017)
```

If you want to play out your fantasy, though, you can specify any bracket
you'd like, as long as it's a character vector of length 64. Once you've done
so, you can use the **sim.bracket** function to play out your own personal
tournament and the **draw.bracket** function to display the outcome.

```{r,dpi=150,fig.height=6,fig.width=8,out.height="500px",out.width="700px"}
set.seed(2017)
outcome = sim.bracket(bracket.empty = bracket.men.2017,
  prob.matrix = prob.matrix)
draw.bracket(bracket.empty = bracket.men.2017, bracket.filled = outcome)
```

Congratulations are in order to Louisville, for winning the set.seed(2017) NCAA
men's basketball tournament!

### Finding a good bracket

That's all well and good for Louisville, but wouldn't you prefer to win
something yourself? Now you can, using the **find.bracket** function. This
function produces a number (*num.candidates*) of candidate brackets and then
evaluates each of them across a number (*num.sims*) of simulations. The larger
you make *num.candidates*  and *num.sims*, the better the bracket will be, but
at the cost of increased computation time. You can choose whether you want the
bracket which maximizes your expected score, expected percentile within your
pool, or even the probability of winning your pool!

You can also customize these results based on the scoring rules and size
(excluding you) of your pool. Below we search for a good bracket to use if we
want to maximize our chances of winning a pool with 30 other people in it,
using the default scoring rules for CBS Sports.

```{r,dpi=150,fig.height=6,fig.width=8,out.height="500px",out.width="700px"}
set.seed(42)
my.bracket = find.bracket(bracket.empty = bracket.men.2017,
  prob.matrix = prob.matrix, num.candidates = 100, num.sims = 1000,
  criterion = "win", pool.size = 30, bonus.round = c(1, 2, 4, 8, 16, 32),
  bonus.seed = rep(0, 16), bonus.combine = "add")
draw.bracket(bracket.empty = bracket.men.2017, bracket.filled = my.bracket)
```

### Testing your bracket

So you're picking Gonzaga? Let's see how well we can expect this bracket to
do, using the **test.bracket** function. This function simulates your pool to
determine your expected score, expected percentile and probability of winning.

```{r}
set.seed(8675309)
test = test.bracket(bracket.empty = bracket.men.2017,
  bracket.picks = my.bracket, prob.matrix = prob.matrix, pool.size = 30,
  num.sims = 1000, bonus.round = c(1, 2, 4, 8, 16, 32),
  bonus.seed = rep(0, 16), bonus.combine = "add")
hist(test$score, breaks = 20)
hist(test$percentile, breaks = 20)
mean(test$win)
```

There you have it. You can expect this bracket to win 12.5% of the Groundhog
Day replays of your bracket pool. Here's to the "real" outcome being one of
those 12.5%!
